Resonant Scattering in a Uniform Sphere with Large Optical Depth

### 2021-08-31 ms.tex
Updates to figures

\begin{figure}
    \centering
    \includegraphics{final_residual.pdf}
    \caption{Analytic solution for the outgoing spectrum compared with Monte Carlo at a line center optical depth of $\tau_0 = 1 \times 10^7$. Photons are initialized with $\rm x = 0.0$. In the log-scale plot in the second panel, $|\rm H_{bc}|$ is shown instead of $\rm H_{bc}$. Residuals of each solution to the Monte Carlo data points are shown in the third panel.} 
    \label{fig:sol_mc_residual_0}
\end{figure}

We now compare each of the previously-discussed solutions for surface flux with Monte Carlo simulations. The escape probability $P(x)$ is defined as the total luminosity at the surface divided by the source luminosity, or
\be
P(x) = \frac{16\pi^2R^2H(R, x)\Delta}{L}
\ee
This is defined such that $\int P(x)dx = 1$. Since $H(r, \nu)$ is implicitly written per $d\nu$, a factor of $\Delta$ gives the expression the correct units. In Fig. \ref{fig:sol_mc_residual_0}, the Monte Carlo escape probability as a function of frequency is shown with that of the solutions $H_{\rm d}$, $H_{\rm 0}$, and $H_{\rm 0 + bc} = H_{\rm 0} + H_{\rm bc}$ for an optical depth of $\tau_0 = 10^7$ and photons emitted at line center $\rm x_s = 0$.  The solution with the correct frequency-dependent boundary condition enforced, $H_{\rm 0 + bc}$, has considerably lower residuals to Monte Carlo than the other solutions, especially in the line wing. It is evident that the divergent solution $H_{\rm d}$ fails in the line wings. The $H_{\rm bc}$ term, when added to the solution $H_0$ derived by \citet{2006ApJ...649...14D}, adequately corrects for the apparent excess of flux in the peaks of the spectrum. This peak excess is also present in previous work that used the \citet{1990ApJ...350..216N} solution, such as \citet{2015MNRAS.449.4336S} where the discrepancy is particularly noticeable for line center optical depths of $\tau_0=10^5$ and $10^6$. The boundary term resolves the slight deficit of $H_{\rm0}$ in the line wings, further improving agreement with the numerical result. Note that the residuals to the $H_0$ solution are a close match to the $H_{\rm bc}$ term, since the Monte Carlo represents the ``true'' solution, $H$, and $H_{\rm bc} = H - H_0$.

 \begin{figure}
    \centering
    \includegraphics{tau_threepanel.pdf}
    \caption{Increasing the line center optical depth. Shown here are $\tau_0 = 10^5, 10^6, 10^7$ with the $x$ and $y$ axes scaled by the characteristic frequency diffusion scale $(a\tau_0)^{1/3}$. Photons are initialized with $\rm x = 0.0$ in each case.} 
    \label{fig:sol_mc_tau}
\end{figure}

@@

 \begin{figure}
    \centering
    \includegraphics{xinit_threepanel.pdf}
    \caption{Analytic solution for the outgoing spectrum compared with Monte Carlo at a line center optical depth of $\tau_0 = 1 \times 10^7$. Photons are initialized with $\rm x = 0.0, 6.0, 12.0$ to examine the solutions' accuracy with a changing source frequency. The optical depth at each of these source frequencies is $\tau_s = 10^7, 77.2,$ and $18.6$, respectively.} 
    \label{fig:sol_mc_xinit}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics{xinit.pdf}
    \caption{$H_{\rm bc}$ solutions shown at three line center optical depths, $\tau_0=10^5, 10^7,$ and $10^9$, for photon source frequencies $\sigma_s=0.0, \tau_0, 2\tau_0$. Thicker lines indicate larger optical depth, and darker lines indicate $\sigma_s$ further from line center. The axes are scaled to show that the size of the corrective factor agrees with the predicted $(a\tau_0)^{-1/3}$ scaling.}
    \label{fig:xinit}
\end{figure}

The previous figures show that as $\tau_0$ increases, $H_0$ appears to become more accurate. There is agreement between the Monte Carlo and $H_0$ when $H_{\rm bc}$ is small which motivates a quantitative examination of how the boundary term $H_{\rm bc}$ scales with both $\tau_0$ and $\sigma_s$, shown in Fig. \ref{fig:xinit}. In this graphic, the photon initialization frequency $\sigma_s$ is shifted in integer multiples of $\tau_0$ such that each curve at a fixed $\sigma_s$ will overlap. For clarity, only the $\sigma > 0$ side of the spectrum is shown. From Eq. \ref{eq:hbc_scaling}, it is expected that the size of the correction $H_{\rm bc}$ should become smaller with larger optical depths following a $(a\tau_0)^{-1/3}$ scaling. This factor has been scaled out of the figure such that solutions for different optical depths and the same source frequency $\sigma_s$ should show close agreement in scale on the figure's vertical axis if the relation holds. Indeed, the scaled solutions converge as the optical depth becomes larger, indicating general agreement with the $(a\tau_0)^{1/3}$ scaling. The remaining discrepancy present in the vertical axis for fixed $\sigma_s$ is due to the width of the Doppler core with respect to the peak of the spectrum; at lower depths, this causes the assumptions about the functional form of the line profile to break down. From this, we conclude that the errors introduced by the incorrect separation of variables in \cite{1973MNRAS.162...43H}, \cite{1990ApJ...350..216N}, \cite{2006ApJ...649...14D} and others are proportional to $(a\tau_0)^{1/3}$, where $a$ is the ratio of natural to Doppler broadening and $\tau_0$ is the optical depth at line center. Furthermore, an additional amount of error of order $(a\tau_0)^{1/3}$ is introduced as the photon frequency $\sigma_s$ from line center is increased by a factor of $\tau_0$.

>>>>

\begin{figure}
    \centering
    \includegraphics{final_residual.pdf}
    \caption{Analytic solution for the outgoing spectrum compared with Monte Carlo at a line center optical depth of $\tau_0 = 1 \times 10^7$. Photons are initialized with $\rm x = 0.0$. In the log-scale plot in the second panel, $|\rm H_{bc}|$ is shown instead of $\rm H_{bc}$ to examine the contributions of both positive and negative pieces. Residuals of each solution to the Monte Carlo data points are shown in the third panel.} 
    \label{fig:sol_mc_residual_0}
\end{figure}

We now compare each of the previously-discussed solutions for surface flux with Monte Carlo simulations. The escape probability $P(x)$ is defined as the total luminosity at the surface divided by the source luminosity, or
\be
P(x) = \frac{16\pi^2R^2H(R, x)\Delta}{L}
\ee
This is defined such that $\int P(x)dx = 1$. Since $H(r, \nu)$ is implicitly written per $d\nu$, a factor of $\Delta$ gives the expression the correct units. In Fig. \ref{fig:sol_mc_residual_0}, the Monte Carlo escape probability as a function of frequency is shown with that of the solutions $H_{\rm d}$, $H_{\rm 0}$, and $H_{\rm 0 + bc} = H_{\rm 0} + H_{\rm bc}$ for an optical depth of $\tau_0 = 10^7$ and photons emitted at line center $\rm x_s = 0$.  The solution with the correct frequency-dependent boundary condition enforced, $H_{\rm 0 + bc}$, has considerably lower residuals to Monte Carlo than the other solutions, especially in the line wing. It is evident that the divergent solution $H_{\rm d}$ fails in the line wings. The $H_{\rm bc}$ term is negative at the peak of the spectrum and positive in the line wing such that, when added to the solution $H_0$ derived by \citet{2006ApJ...649...14D}, it adequately corrects for the apparent excess of flux in the peaks of the spectrum. This peak excess is also present in previous work that used the \citet{1990ApJ...350..216N} solution, such as \citet{2015MNRAS.449.4336S} where the discrepancy is particularly noticeable for line center optical depths of $\tau_0=10^5$ and $10^6$. The boundary term resolves the slight deficit of $H_{\rm0}$ in the line wings, further improving agreement with the numerical result. Note that the residuals to the $H_0$ solution are a close match to the $H_{\rm bc}$ term, since the Monte Carlo represents the ``true'' solution, $H$, and $H_{\rm bc} = H - H_0$.

 \begin{figure}
    \centering
    \includegraphics{tau_threepanel.pdf}
    \caption{The same spectrum as in Fig. \ref{fig:sol_mc_residual_0}, but for three different line center optical depths $\tau_0 = 10^5, 10^6, 10^7$ with the $x$ and $y$ axes scaled by the characteristic frequency diffusion scale $(a\tau_0)^{1/3}$. Photons are initialized with $\rm x = 0.0$ in each case.} 
    \label{fig:sol_mc_tau}
\end{figure}

@@

 \begin{figure}
    \centering
    \includegraphics{xinit_threepanel.pdf}
    \caption{Analytic solution for the outgoing spectrum compared with Monte Carlo at a line center optical depth of $\tau_0 = 1 \times 10^7$. Photons are initialized with $\rm x = 0.0, 6.0, 12.0$ to examine the solutions' accuracy with a changing source frequency. The optical depth at each of these source frequencies is $\tau_s = 10^7, 77,$ and $19$, respectively.} 
    \label{fig:sol_mc_xinit}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{xinit.pdf}
    \caption{$H_{\rm bc}$ solutions shown at three line center optical depths, $\tau_0=10^5, 10^7,$ and $10^9$, for photon source frequencies $\sigma_s=0.0, \tau_0, 2\tau_0$. Thicker lines indicate larger optical depth, and darker lines indicate $\sigma_s$ further from line center. The axes are scaled to show that the size of the corrective factor agrees with the predicted $(a\tau_0)^{-1/3}$ scaling.}
    \label{fig:xinit}
\end{figure}

The previous figures show that as $\tau_0$ increases, $H_0$ appears to become more accurate. There is agreement between the Monte Carlo and $H_0$ when $H_{\rm bc}$ is small which motivates a quantitative examination of how the boundary term $H_{\rm bc}$ scales with both $\tau_0$ and $\sigma_s$, shown in Fig. \ref{fig:xinit}. In this graphic, the photon initialization frequency $\sigma_s$ is shifted in integer multiples of $\tau_0$ such that $\sigma_s$ falls near the peak of the spectrum for each $\tau_0$. For clarity, only the $\sigma > 0$ side of the spectrum is shown. From Eq. \ref{eq:hbc_scaling}, it is expected that the size of the correction $H_{\rm bc}$ should become smaller with larger optical depths following a $(a\tau_0)^{-1/3}$ scaling. This factor has been scaled out of the figure such that solutions for different optical depths and the same source frequency $\sigma_s$ should show close agreement in scale on the figure's vertical axis if the relation holds. Indeed, the scaled solutions converge as the optical depth becomes larger, indicating general agreement with the $(a\tau_0)^{1/3}$ scaling. The remaining discrepancy present in the vertical axis for fixed $\sigma_s$ is due to the width of the Doppler core with respect to the peak of the spectrum; at lower depths, this causes the assumptions about the functional form of the line profile to break down. From this, we conclude that the errors introduced by the incorrect separation of variables in \cite{1973MNRAS.162...43H}, \cite{1990ApJ...350..216N}, \cite{2006ApJ...649...14D} and others are proportional to $(a\tau_0)^{1/3}$, where $a$ is the ratio of natural to Doppler broadening and $\tau_0$ is the optical depth at line center. Furthermore, an additional amount of error of order $(a\tau_0)^{1/3}$ is introduced as the photon frequency $\sigma_s$ from line center is increased by a factor of $\tau_0$.
